---
title: "p8105_hw3_yy3439"
output: github_document
date: "2023-10-10"
---

## Problem 1 (All Parts of Problem 1 are from the Solution Provided by Professor Goldsmith, only here for completion)

```{r}
library(p8105.datasets)
library(tidyverse)
library(ggridges)
library(patchwork)
library(ggplot2)
data("instacart")
```

```{r}
instacart = 
  instacart |> 
  as_tibble()
```
```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2


```{r}
library(p8105.datasets)
data("brfss_smart2010")

```


#### Tidying-up Data
```{r}

brfss_smart2010 = 
  brfss_smart2010 |>
  janitor::clean_names()|> #format data to appropriate variable names 
  filter(topic == "Overall Health")|> #focus only on the "Overall Health"
  filter(response %in% c("Excellent","Very Good", "Good", "Fair", "Poor"))|> #include only responses from excellent to poor
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very Good", "Excellent"))) #relevel the responses as a factor from "Poor" to "Excellent"
  
 
 
```

#### 2002 Unique Obervation States

```{r}
states_2002<-brfss_smart2010 |>
  separate(locationdesc, into = c('state', 'location', sep = '-'))|>
  filter(year =='2002')|>
  group_by(state)|>
  summarise(distinct_location = n_distinct(location))|>
  filter(distinct_location > 6)|>
  knitr::kable()

states_2002
```


#### 2010 Unique Obervation States

```{r}
states_2010 <-brfss_smart2010 |>
  separate(locationdesc, into = c('state', 'location', sep = '-'))|>
  filter(year =="2010")|>
  group_by(state)|>
  distinct(location)|>
  summarise(distinct_location = n())|>
  filter(distinct_location > 6)|>
  knitr::kable()

states_2010
```
As shown in the above table, we have 5 states that have 7 or more distinct locations： FL, MA, NC, NJ, and PA. Similarly, we have 14 states that have 7 or more distinct locations: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA.

#### Data_value


```{r}
avrage_brfss <- brfss_smart2010|>
  filter(response == "Excellent")|>
  group_by(year, locationabbr)|>
  select(year,  state = locationabbr, response, data_value)|>
   summarise(
    avg_data_value = mean(data_value, na.rm = TRUE))
  
 avrage_brfss
  
```
#### Spaghetti Plot 


```{r}
avg_excellent = ggplot(avrage_brfss, aes(x = year, y = avg_data_value, group = state))+
  geom_line(aes(color = state))+
  theme_minimal()+
  labs(title = "Average Value of Excellent Responses Over Years", x = "Year", y = "Average Value")
ggsave("average value of excellent.png", avg_excellent)
avg_excellent
```

#### Two Panel Plot

```{r}
panel_brfss = brfss_smart2010|>
  filter(year%in% c(2006, 2010))|>
  filter(locationabbr == "NY")

data_value_dist = ggplot(panel_brfss, aes(x = response, y = data_value, fill = response))+
  geom_boxplot()+
  facet_grid(~year)+
  labs(
    title = "Data Value Distributions for the Year of 2006 and 2010",
    x = "Response",
    y = "Data Value"
  )
  
ggsave("data value distribution.png",data_value_dist )
data_value_dist
```



### Problem 3


#### Load Data
```{r}
#Load Data

nhanes_covar <- read_csv("nhanes_covar.csv", skip = 4)
nhanes_accel <- read_csv("nhanes_accel.csv")
```

#### Organize the `covar` Dataset

```{r}
#Organize the `covar` Dataset
covar_clean = nhanes_covar|>
  janitor::clean_names()|>
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ "male",
        2 ~ "female"),
    sex = as.factor(sex)
      )|> #recode the numerical sex variable to factor variables;this code was borrowed from lecture notes"Tidy Data"
  mutate(
    education = 
      case_match(
        education,
        1 ~ "Less than high school",
        2 ~ "High school equivalent",
        3 ~ "More than high school"),
    education = as.factor(education)
      )|>
  filter(age >= 21)|> #filter out only participants aged 21 or 21+
  drop_na() #drop NA values
  
  

```

#### Organize the `accel` Dataset

```{r}
#Organize the `accel` Dataset

accel_clean = nhanes_accel|>
  janitor::clean_names()|>
  drop_na()
  
```

#### Joinging Data set

```{r}
# Joinging Data set
joined_dataset <- left_join(covar_clean, accel_clean, by = "seqn") 
```

The final merged data set contains all variables from the `covar` and `accel` datasets and joined by the variable `seqn`, which appears in both datasets. 

#### Table 1: Number of Male and Female in each of the Education Category

```{r}
edu_gender <- joined_dataset|>
  group_by(sex, education)|>
  drop_na()|>
  summarise(count = n(), .groups = "drop")|>
  pivot_wider(
    names_from = sex,
    values_from = count
  )|>
  knitr::kable(caption = "Number of Male and Female in Each Education Level")
```
As shown in the table, the total number of female participants who have less than and more than high school degrees is larger than the total number of male participants. In contrast, the total number of male participants who have high school degrees is larger than the female participants. 


#### Visualization of the Age Distribution for Male and Female Participants


```{r}
joined_clean_data <- joined_dataset|>drop_na()
age_dist = ggplot(joined_clean_data, 
       aes(x = education, y = age, color = sex)) +
       geom_boxplot()+
       labs(
         title = "Age Distribution for Education Level of Male and Female Participants",
         x = "Education Level",
         y = "Age"
       )+
  theme_minimal()
ggsave("age_disttribution.png", age_dist )  
age_dist
```
As shown in the boxplot, the medians of age distribution in which females who have high school or less than high school degrees are larger than the male participants. Specifically, people who aged below 50s tend to have higher than high school degrees, and the age distribution in which male and female participants having less than high school degrees are similar. 



#### Total Activity Time per Participant
```{r}
activity_par <- joined_clean_data|>
  group_by(seqn)|>
  pivot_longer(
    min1:min1440,
    names_to = "time",
    values_to = "activity"
  )|>
  summarise(total_activity_time = sum(activity))
```


#### Visualization of the Total Activity Time per Participant

```{r}
total_activity = inner_join(joined_clean_data, activity_par, by = "seqn")|>
  relocate(seqn, sex, age, bmi, education,total_activity_time ) #binding the two dataset and relocate so we can see the total activity time more easily 

```
```{r}
#Visualization
activity_plot = ggplot(total_activity, aes(x = age, y = total_activity_time, fill = sex))+
  geom_point()+
  geom_smooth()+
  labs(
    title = "Total Activity Time for Each Participant",
    x = "Age",
    y = "Total Activity Time"
  )+
  facet_grid(.~education)
ggsave("total_activity_time.png", activity_plot)
activity_plot
```
##### Discussion
As shown in the plot above, male with high school equivalaent and more than high school degrees generally have more activity time compared to their female counterparts. However, female with less than high school degrees have more activity time on average compared to their male counterparts across the age group. For all male and female, people who are younger than 40 years old have the most activity time whereas the activity time dropped eminently for people aged 60 or above. 


#### Three-Panel Plot

```{r}
minute = joined_clean_data|>
  pivot_longer(
    min1:min1440,
    names_to = "time",
    values_to = "activity"
  )|>
separate(time, into = c("time", "minute"), sep = 3)|>
  mutate(minute = as.numeric(minute))
```




```{r}
minute_plot = ggplot(minute, aes(x = minute, y = activity, color = sex))+
  geom_smooth(se = FALSE)+
  facet_grid(.~education)+
  labs(
    title = "24 Hour Activity by Sex and Education Level",
    x = "Time in Minute",
    y = "Activity Time"
  )+
  scale_x_continuous(
    breaks = c(0, 240, 480, 720, 960, 1200, 1440)
  )+
  theme(legend.position = "bottom")
minute_plot
```
```{r}
ggsave("24_Hour_Activity_Time.png", minute_plot)
```


##### Discussion
Based on the three-panel graph, we can see that the activity time dropped significantly from 0 to min240. Then, the activity time increased sharply from min240 to min600 for all male and female participants. The activity time for male and female participants with less than high school degrees would decrease directly once past the min600 threshold However, for male and female participants with high school equivalent or more than high school degrees, the activity time would fluctuate from min600 to min1200 and then decreased significantly from min1200. In addition, female participants tend to have more activity time on average compared to the male participants across all education level. The three panel graph indicates that people with higher education levels tend to be more stable and less likely to be distracted compared to people with lower education levels. 




