---
title: "p8105_hw3_yy3439"
output: github_document
date: "2023-10-10"
---

## Problem 1

```{r}
library(p8105.datasets)
library(tidyverse)
library(ggridges)
library(patchwork)
library(ggplot2)
data("instacart")
```

```{r}
instacart = 
  instacart |> 
  as_tibble()
```




## Problem 2


```{r}
library(p8105.datasets)
data("brfss_smart2010")

```


#### Tidying-up Data
```{r}

brfss_smart2010 = 
  brfss_smart2010 |>
  janitor::clean_names()|> #format data to appropriate variable names 
  filter(topic == "Overall Health")|> #focus only on the "Overall Health"
  filter(response %in% c("Excellent","Very Good", "Good", "Fair", "Poor"))|> #include only responses from excellent to poor
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very Good", "Excellent"))) #relevel the responses as a factor from "Poor" to "Excellent"
  
 
 
```

#### 2002 Unique Obervation States

```{r}
states_2002<-brfss_smart2010 |>
  separate(locationdesc, into = c('state', 'location', sep = '-'))|>
  filter(year =="2002")|>
  group_by(state)|>
  distinct(location)|>
  summarise(distinct_location = n())|>
  filter(distinct_location > 6)

states_2002
```
As shown in the above tibble, we have 5 states that have 7 or more distinct locations： FL, MA, NC, NJ, and PA. Similarly, as shown in the tibble below, we have 14 states that have 7 or more distinct locations: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA.


#### 2010 Unique Obervation States

```{r}
states_2010 <-brfss_smart2010 |>
  separate(locationdesc, into = c('state', 'location', sep = '-'))|>
  filter(year =="2010")|>
  group_by(state)|>
  distinct(location)|>
  summarise(distinct_location = n())|>
  filter(distinct_location > 6)

states_2010
```


#### Data_value


```{r}
avrage_brfss <- brfss_smart2010|>
  filter(response == "Excellent")|>
  group_by(year, locationabbr)|>
  select(year,  state = locationabbr, response, data_value)|>
   summarise(
    avg_data_value = mean(data_value, na.rm = TRUE))
  
 avrage_brfss
  
```
#### Spaghetti Plot 


```{r}
ggplot(avrage_brfss, aes(x = year, y = avg_data_value, group = state))+
  geom_line(aes(color = state))+
  theme_minimal()+
  labs(title = "Average Value of Excellent Responses Over Years", x = "Year", y = "Average Value")
```

#### Two Panel Plot

```{r}
panel_brfss = brfss_smart2010|>
  filter(year%in% c(2006, 2010))|>
  filter(locationabbr == "NY")

ggplot(panel_brfss, aes(x = response, y = data_value, fill = response))+
  geom_boxplot()+
  facet_grid(~year)+
  labs(
    title = "Data Value Distributions for the Year of 2006 and 2010",
    x = "Response",
    y = "Data Value"
  )
  
  
  
```



### Problem 3


```{r}
#Load Data

nhanes_covar <- read_csv("nhanes_covar.csv", skip = 4)
nhanes_accel <- read_csv("nhanes_accel.csv")
```

```{r}
#Organize the `covar` Dataset
covar_clean = nhanes_covar|>
  janitor::clean_names()|>
  mutate(
    sex = 
      case_match(
        sex,
        1 ~ "male",
        2 ~ "female"),
    sex = as.factor(sex)
      )|> #recode the numerical sex variable to factor variables;this code was borrowed from lecture notes"Tidy Data"
  mutate(
    education = 
      case_match(
        education,
        1 ~ "Less than high school",
        2 ~ "High school equivalent",
        3 ~ "More than high school"),
    education = as.factor(education)
      )|>
  filter(age >= 21)|> #filter out only participants aged 21 or 21+
  drop_na() #drop NA values
  
  

```

```{r}
#Organize the `accel` Dataset

accel_clean = nhanes_accel|>
  janitor::clean_names()|>
  drop_na()
  
```

```{r}
# Joinging Data set
joined_dataset <- left_join(covar_clean, accel_clean, by = "seqn") 
```

The final merged data set contains all variables from the `covar` and `accel` datasets and joined by the variable `seqn`, which appears in both datasets. 

#### Table 1: Number of Male and Female in each of the Education Category

```{r}
edu_gender <- joined_dataset|>
  group_by(sex, education)|>
  drop_na()|>
  summarise(count = n(), .groups = "drop")|>
  pivot_wider(
    names_from = sex,
    values_from = count
  )|>
  knitr::kable(caption = "Number of Male and Female in Each Education Level")
```
As shown in the table, the total number of female participants who have less than and more than high school degrees is larger than the total number of male participants. In contrast, the total number of male participants who have high school degrees is larger than the female participants. 

#### Visualization of the Age Distribution for Male and Female Participants


```{r}
joined_clean_data <- joined_dataset|>drop_na()
ggplot(joined_clean_data, 
       aes(x = education, y = age, color = sex)) +
       geom_boxplot()+
       labs(
         title = "Age Distribution for Education Level of Male and Female Participants",
         x = "Education Level",
         y = "Age"
       )+
  theme_minimal()
  
```
As shown in the boxplot, the medians of age distribution in which females who have high school or less than high school degrees are larger than the male participants. Specifically, people who aged below 50s tend to have higher than high school degrees, and the age distribution in which male and female participants having less than high school degrees are similar. 



#### Total Activity Time per Participant
```{r}
activity_par <- joined_clean_data|>
  group_by(seqn)|>
  pivot_longer(
    min1:min1440,
    names_to = "time",
    values_to = "activity"
  )|>
  summarise(total_activity_time = sum(activity))
```


#### Visualization of the Total Activity Time per Participant

```{r}
total_activity = inner_join(joined_clean_data, activity_par, by = "seqn")|>
  relocate(seqn, sex, age, bmi, education,total_activity_time ) #binding the two dataset and relocate so we can see the total activity time more easily 

```
```{r}
#Visualization
ggplot(total_activity, aes(x = age, y = total_activity_time, fill = sex))+
  geom_point()+
  geom_smooth()+
  labs(
    title = "Total Activity Time for Each Participant",
    x = "Age",
    y = "Total Activity Time"
  )+
  facet_grid(.~education)
```


